---
title: "Assignment B-1"
author: "Clayton"
date: "2023-10-21"
output: html_document
---

# The Function

This is the dot product. Given two vectors $x$ and $y$ of same dimension, each item of $x$ is multiplied by the item of $y$ in the same position. Then all the products are multiplied. This is given the notation $x \dot y$. For example, if $x=[1, -1], y=[2, -1]$, we have $x \dot y = 1*2 + -1*-1=3$

```{r}
#' Computes the dot product of x and y. Throws an error if either vector are non-numeric or contain NA, or if their lengths mismatch.
#' 
#' The naming convention of x and y are traditional to represent vectors
#' 
#' @param x A numeric vector. 
#' @param y A numeric vector.
#'
#' @return The result of the dot product operation.
dot_product <- function(x, y){
  # throw error of x or y are not numeric
  stopifnot(is.numeric(x) && is.numeric(y))
  # throw error if x and y do not have matching dimensions.
  stopifnot(length(x) == length(y))
  # throw error if either x or y is NA.
  stopifnot(!(anyNA(x) || anyNA(y)))
  
  sum(x*y)
}
```

# Examples

For the first example, we will look at the example $x=[1, -1], y=[2, -1]$ from above. This should be equal to $3$.

```{r}
# initialize vectors
x = c(1, -1)
y = c(2, -1)

#output
value = dot_product(x, y)
paste("The dot product of x and y is:", value)
```

One common application of computing the dot product is to find the length of the projection of one vector onto another. The formula for that is $$\text{proj}_v(u) = (u\dot v)*v$$ where $v$ is a unit vector (has length $1$).

```{r}
# this is a unit vector
v = c(3, 4) / 5
u = c(2, 3)

# get values
length_of_projection = dot_product(u, v)
projection = length_of_projection*v

# output
paste("The length of the projection is equal to:", length_of_projection)
paste("The u vector projects onto v at:", paste(projection, collapse = ', '))
```

# Tests

Below we test that the function works on valid errors, and that it throws the expected errors.

```{r}
library(testthat)
```

```{r}
test_that("Works on valid inputs", {
  expect_equal(10, dot_product(c(4, 6), c(1, 1)))
  expect_equal(0, dot_product(c(0, 0), c(1, 1)))
  expect_equal(2, dot_product(c(1, -1, -3), c(3, -2, 1)))
  expect_equal(.02, dot_product(c(.1, -.1, -.3), c(.3, -.2, .1)))
})
```

```{r}
test_that("Works on scalar values", {
  expect_equal(3, dot_product(c(1), c(3)))
  expect_equal(3, dot_product(1, 3))
})
```

```{r}
test_that("Non-numeric produces error", {
  expect_error(dot_product(c('1'), c(3)))
  expect_error(dot_product(c(1, 4), c(3, '0')))
})
```

```{r}
test_that("Length Mismatch", {
  expect_error(dot_product(c(1), c(3, 4)))
  expect_error(dot_product(c(1, 4, -2), c(1, 0)))
})
```

```{r}
test_that("Test on NA values", {
  expect_error(dot_product(NA, c(3, 4)))
  expect_error(dot_product(c(1, 4), c(1, NA)))
})
```
